---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
```{r echo=FALSE}
options(scipen=1,digits=12)
```

## Loading and preprocessing the data

```{r load data}
#load data and necessary libraries
library(dplyr)
library(ggplot2)
AcData <- read.csv("activity.csv", na.strings="NA")
```

## What is mean total number of steps taken per day?

```{r Total Steps}
#total number of steps per day:
T_steps <- aggregate(steps ~ date, data=AcData, FUN=sum) 
#histrogram of total number of steps per day:
hist(T_steps$steps,  xlab="Total number of steps taken per day", main="", breaks=10)
#output summary
summary(T_steps$steps)
```

The mean of the total number of steps per day is : `r mean(T_steps$steps)`

The median of the total number of steps per day is : `r median(T_steps$steps)`

## What is the average daily activity pattern?

```{r Daily Activity Pattern}
#calculate the mean number of steps taken per interval
step_int <- aggregate(steps ~ interval, data=AcData, FUN=mean)
#time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
qplot(step_int$interval,step_int$steps, geom="path",xlab="Interval",ylab="Avg number of steps")
summary(step_int$steps)
```

The 5-minute interval containing the maximum number of steps, averaged across all days in the data set is: `r step_int$interval[which.max(step_int$steps)]`. The maximun number of averaged steps is: `r max(step_int$steps)`.

## Imputing missing values

```{r}
MV <- sum(is.na(AcData$steps))
```

The number of missing values is: `r MV`.

We replace all the missing values with the mean for that 5-minute interval:

```{r Missing Values}
head(AcData)
intmeans <- rep(step_int$steps, 61)
#create a new data set with missing values filled in:
AcData_new <- mutate(AcData, steps = replace(AcData$steps, is.na(AcData$steps),intmeans[is.na(AcData$steps)]))
head(AcData_new)

#total number of steps per day in new data set:
T_steps_new <- aggregate(steps ~ date, data=AcData_new, FUN=sum) 
#histrogram of total number of steps per day:
hist(T_steps_new$steps,  xlab="Total number of steps taken per day", main="", breaks=10)
#output summary
summary(T_steps_new$steps)
```

The mean of the total number of steps per day is : `r mean(T_steps_new$steps)`

The median of the total number of steps per day is : `r median(T_steps_new$steps)`

Imputing missing values using the approach described above does not change the mean values, however it does slightly alter the median value by: `r ((median(T_steps_new$steps) - median(T_steps$steps))/median(T_steps$steps))*100` %.


## Are there differences in activity patterns between weekdays and weekends?

```{r Days of the Week}
#modify date format in data set with filled-in missing values
AcData_new$date <- as.character(AcData_new$date)
AcData_new$date <- as.Date(AcData_new$date)
#add new variable for weekday name:
AcData_new <- mutate(AcData_new, weekday=weekdays(AcData_new$date, abbreviate=FALSE))
#create factor variable: day, indicating weekend or weekday
AcData_new$day <- ifelse(AcData_new$weekday %in% c("Saturday", "Sunday"), "weekend", "weekday")
AcData_new$day <- as.factor(AcData_new$day)
str(AcData_new)

#calculate mean steps per int: weekday & weekend
step_int_weekday <- aggregate(steps ~ interval + day, data=AcData_new, FUN=mean) 
#plot:
g <- ggplot(step_int_weekday, aes(interval,steps))
g + geom_line() + facet_grid(day ~ ., as.table=FALSE)
```

